---
title: "index.Rmd"
author: "Xiaotong(Katherine) Zhang, Yenmy Vo, Natalie Hinds, Kasper Li"
date: "November 14, 2022"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyr")
library("knitr")
library("dplyr")
source("https://raw.githubusercontent.com/info201b-au2022/project-itsyenmyvo22/main/source/summary_information.R")
source("https://raw.githubusercontent.com/info201b-au2022/project-itsyenmyvo22/main/source/summary_table.R")
source("https://raw.githubusercontent.com/info201b-au2022/project-itsyenmyvo22/main/data/poverty_nkilled_graph.R")
source("https://raw.githubusercontent.com/info201b-au2022/project-itsyenmyvo22/main/the%20number%20of%20gun%20violence%20%26%20injured.R")
source("https://raw.githubusercontent.com/info201b-au2022/project-itsyenmyvo22/main/Gun%20universal%20background.R")
```


# **Gun Violence in the US: Project Proposal**


## Code Name: waterguns

## _Authors:_
**Natalie Hinds (Nhinds2@uw.edu)
 Kasper Li(jiaxul9@uw.edu)
 Katherine Zhang (Xiaotz7@uw.edu)
 Yenmy Vo (yenmyvo@uw.edu)**

### INFO-201: Technical Foundations of Informatics - The Information School - University of Washington

November 2022

## Summary Information

```{r, echo=FALSE, message = F, include=FALSE}
info <- summary_info
info$deaths_sales_18
info$deaths_sales_21
info$max_deaths_state
info$highest_rate_state
info$highest_deaths_one_day
```

We use datasets that are specific to incidents of gun violence, such as the state where the violence occurred, the date it occurred, how many casualties there were, how many areas sold guns to 18+, how many areas sold guns to 21+, and local poverty rates. The dataset we chose includes U.S. gun violence incidents from 2013 to 2018. To demonstrate the main purpose of what our project is studying, we chose to calculate five variables through the dataset that are representative of our research questions. These are the states in the dataset with the highest number of deaths due to gun violence, the states with the highest poverty rates, the highest number of deaths in a day, the number of deaths in areas where guns are sold to people over 18 years old, and the number of deaths in areas where guns are sold to people over 21 years old. The results obtained were **`r info$deaths_sales_18`** deaths in areas where guns were sold to people over the age of 18 and **`r info$deaths_sales_21`** deaths in areas where guns were sold to people over the age of 21. The data set showed **`r info$highest_deaths_one_day`** gun violence deaths in a day, and the state with the highest number of deaths due to gun violence was **`r info$max_deaths_state`**. In addition, the state with the highest poverty rate was **`r info$highest_rate_state`**.


### Summary Table

```{r, echo=FALSE, eval=TRUE}
summary_table
```

* This table shows the five values we chose to calculate. As you can visually see from the table, the number of gun violence deaths in areas where guns are sold to those over 18 is much higher than the number of deaths in areas where guns are sold to those over 21. Also, it can be concluded from the maximum number of deaths in one day due to gun violence that gun violence is a very serious cause for concern.


## Charts

### Chart1: Poverty

```{r, echo=FALSE, eval=TRUE}
plot(trial)
```

* This chart was included to show the relationship between gun violence and poverty levels. The chart reveals that states with higher poverty levels tend to have more gun deaths. New Hampshire, North Dakota, and Wyoming are some of the states with the highest percentage of people above poverty. These are also the states with some of the lowest gun deaths. There's a trend that the midwest has a low number of gun deaths while the eastern states as well as the states lining the southern border have high gun deaths. 

### Chart2: Universal Background Checks and Gun Deaths.

```{r, echo=FALSE, eval=TRUE}
plot(universal)
```

* I chose to include the chart to show the amount of deaths and the effect that background checks have on deaths in each state. The information reveals that only 9 states have implemented gun control laws with universal background checks. The chart also reveals that many states without universal background checks have a significant proportion of deaths compared to states with universal background deaths, with an exception of California. The data visualization does not account for population density, as states with rural populations tend to have a lower amount of gun violence. 


### Chart3: The number of gun violence & injured

```{r, echo=FALSE, eval=TRUE}
plot(Killed_injured)
```

* This chart shows the number of people killed and injured by shootings across the United States over the years. Since there are many variables in the original data, I simplified it to monthly data. The chart shows that the number of injuries and deaths from 2013 to 2014 is not very large. But since 2014, the number has soared, and the trend has been increasing year by year. For example, in 2014, the maximum number of people injured in one month was about 2,500, but in 2016, the maximum number of people injured in one month rose to 3,000. Therefore, we can conclude from this data visualization that gun violence is gradually getting worse.
